---
title: "Untitled"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tictoc)
library(ggmap)
library(skimr)
library(lubridate)
library(forcats)
library(leaflet)
```

#Downloading data
```{r, include=FALSE}
#year 2017
URL <- "https://s3.amazonaws.com/fordgobike-data/2017-fordgobike-tripdata.csv"
download.file(URL, destfile = "./data/2017-fordgobike-tripdata.csv", method="curl")

#year 2018

for (i in 1:7) {
URL <- paste0("https://s3.amazonaws.com/fordgobike-data/20180",i,"-fordgobike-tripdata.csv.zip")
download.file(URL, destfile = paste0("./data/20180",i,"-fordgobike-tripdata.csv.zip"), method="curl")
}

#year 2019

for (i in 1:4) {
URL <- paste0("https://s3.amazonaws.com/fordgobike-data/20190",i,"-fordgobike-tripdata.csv.zip")
download.file(URL, destfile = paste0("./data/20190",i,"-fordgobike-tripdata.csv.zip"), method="curl")
}

```
#Unzipping files
```{r, include=FALSE}
#year2018
unzip("./data/201801-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201802-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201803-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201804-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201805-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201806-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201807-fordgobike-tripdata.csv.zip",exdir="./data")

#year2019
unzip("./data/201901-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201902-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201903-fordgobike-tripdata.csv.zip",exdir="./data")
unzip("./data/201904-fordgobike-tripdata.csv.zip",exdir="./data")


```

#removing zipped files
```{r, include=FALSE}
#year 2018
fn <- "./data/201801-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201802-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201803-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201804-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201805-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201806-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201807-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

#year 2019
fn <- "./data/201901-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201902-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201903-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)

fn <- "./data/201904-fordgobike-tripdata.csv.zip"
if (file.exists(fn)) file.remove(fn)


```
#Reading csv files
```{r, include=FALSE}
fordgobike2017 <- read_csv(file="./data/2017-fordgobike-tripdata.csv")
fordgobike201801 <- read_csv(file="./data/201801-fordgobike-tripdata.csv")
fordgobike201802 <- read_csv(file="./data/201802-fordgobike-tripdata.csv")
fordgobike201803 <- read_csv(file="./data/201803-fordgobike-tripdata.csv")
fordgobike201804 <- read_csv(file="./data/201804-fordgobike-tripdata.csv")
fordgobike201805 <- read_csv(file="./data/201805-fordgobike-tripdata.csv")
fordgobike201806 <- read_csv(file="./data/201806-fordgobike-tripdata.csv")
fordgobike201807 <- read_csv(file="./data/201807-fordgobike-tripdata.csv")
fordgobike201901 <- read_csv(file="./data/201901-fordgobike-tripdata.csv")
fordgobike201902 <- read_csv(file="./data/201902-fordgobike-tripdata.csv")
fordgobike201903 <- read_csv(file="./data/201903-fordgobike-tripdata.csv")
fordgobike201904 <- read_csv(file="./data/201904-fordgobike-tripdata.csv")

```
# changing variable type before binding data set
```{r, include=FALSE}
#changes in year2018
fordgobike201806 <- fordgobike201806 %>%
        mutate(start_station_id = as.integer(start_station_id),
               end_station_id= as.integer(end_station_id) ) 
fordgobike201807 <- fordgobike201807 %>%
        mutate(start_station_id = as.integer(start_station_id),
               end_station_id= as.integer(end_station_id) )

#changes in year 2019
fordgobike201902 <- fordgobike201902 %>%
        mutate(start_station_id = as.integer(start_station_id),
               end_station_id= as.integer(end_station_id) )

fordgobike201903 <- fordgobike201903 %>%
        mutate(start_station_id = as.integer(start_station_id),
               end_station_id= as.integer(end_station_id) )


```


#Combining data frame year wise
```{r, include=FALSE}
#year2018
fordgobike2018 <- bind_rows(fordgobike201801, fordgobike201802, fordgobike201803, fordgobike201804,
                            fordgobike201805, fordgobike201806, fordgobike201807)

#year2019
fordgobike2019 <- bind_rows(fordgobike201901, fordgobike201902, fordgobike201903, fordgobike201904)

# year 2017,2018,2019
fordgobike_all<-bind_rows(fordgobike2017,fordgobike2018,fordgobike2019)
dim(fordgobike_all)
str(fordgobike_all)

```

#creating separate columns for days, months and years
```{r, include=FALSE}
fordgobike_all <- fordgobike_all %>% mutate(year=year(start_time), month=month(start_time), day=day(start_time) ) 
fordgobike_all%>%group_by(year)%>%summarize(n=n())
head(fordgobike_all)
```
#Before creating map, lets see which station has the highest number of pick ups
#This flexdashboard is trying to show busiest station in chart(a), busiest month in chart(b) and the year wise usage of fordgobikes by years in chart(c).
```{r, include=FALSE}
fordgobike_all%>%
  select(start_station_id,start_station_latitude,start_station_longitude)%>%
  group_by(start_station_id)%>%
  summarize(n=n())%>%
  arrange(desc(n))

#San Francisco Ferry Building (Harry Bridges Plaza): lat=37.79539	 and long =-122.3942
fordgobike_all%>%filter(start_station_id==15)%>%select(start_station_latitude,start_station_longitude)
```


Column {data-width=650}
-----------------------------------------------------------------------

### Busiest station for fordgobike

```{r}
mostfamous_station <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=-122.3942, lat=37.79539, popup="Busiest station for fordgobike")  
mostfamous_station
```

Column {data-width=350}
-----------------------------------------------------------------------

### Hisotgram of the Duration of Fordgobike Trips for each Gender

```{r}
graph2=fordgobike19 %>% ggplot(aes(log(x=duration_sec), y=..density..,fill=member_gender)) + 
  geom_histogram() + 
  geom_density(aes(y=..density..)) +
  facet_grid(member_gender ~ .)+ggtitle("")
ggplotly(graph2)
```

###yearly data showing usage of fordgobikes
```{r}
fordgobike_all %>% ggplot(aes(x=month)) + geom_bar() + facet_grid(year ~ .)+ggtitle("yearly data showing usage of fordgobikes")
```


